---
title: "R Notebook"
output: html_notebook
---
#Recommendation
Recommendation matrix loko at the history of data, if all have seen movie A it cannot recomment it. If 2 out fo 4 have seen movie B, it can rec to see Movie B to the other two, 

  S E T M
Y 1 1 1 0
C 0 1 1 0 
M 0 0 1 1     
K 1 1 1 1

Y + C = 2
Y + M = 1
Y + K = 3

  Y C M K
Y 1 2 1 3
C 2 1 . .
M 1 . 1 .
K 3 . . 1



-Build diff candidate models, write a test thingy to how good they are..
-Recommend and compare with om den blev kopt,
-Data have to be in a matrix, all values the same datatype.
-
- Create a Rating matrix, cons: extream data, very satisfied or Absolutely not -- > need to normalize using std from mean.
- remove those who did not like the movie, keep movies that people recommended little, up to very much. remove the contiuous values


## example 1
```{r}
library(recommenderlab)
library(tidyverse)

data(MovieLense) # good and bad movies 
image(MovieLense)
model <- Recommender(MovieLense, method = "POPULAR")
#as(predict(model, MovieLense[101], n=10))  # make prediction use prediction engine, it wants reccomendation, convert the object, 


#expand.grid()


```






## example 2, jokes

-Popular jokes that he has not seen is recommendended.
-How would the person rates jokes in relation to popularity estimated, high or low ranked
-Users actual rating the user gave and compared to predicted, look were errors are.
-Evaluation scheme under which our model will work
- Train-test-split: how many random users we shuld keep...
- Build the evaluation scheme, to preform useful stuff...
- Predict data, give ratings, give back as ratings to see what they look like
- calc accuracy to compare against unknown data
- create a crossvalidation matrix
- evaluate function will build a model for us, use the evaluate, build it x nr of times.
RMSE = rms , MSE= mean square error, MAE =mean absolute error
- Can get the average of these values, the blended preformance of our model, based on 10 diff samples.
-Create an evaluation scheme, ..operate on rating data...
-Build a single model, based on rating solution.



```{r}

data()

```

