---
title: "R Notebook"
output: word_document  # pdf_documents, html_notebook, html_document
---

# Abuot R
- R came from S
- R works in memory
    + Microsoft has out-of-memory processing capabilities
    + Theres some other potential options
- R is quirky AF
- R is keysensitive
- There's a package for that!


# insta
```{r eval=FALSE}
install.packages('reticulate')
```

```{python}
print('hello')
```

```{r}
library("tidyverse")
```


```{r}
tidyverse_packages()
```


# Vectors
1d objects, c menas combine
```{r}
c("red","blue", "green")

colors <- c("red","blue", "green")
colors2 <- c("purpule", "orange")

1:10

nums <- 1:10
nums

about <- c(firstname="Steph", lastname = "Andersson")
about


```



```{r}
colors
colors[1]
colors[-1]
colors[c(1,3)]
colors[1:2]
colors[-c(1,3)]  # excludes 1 and 3, shows other values

runif(n=1,min=1,max=3)   #randomly generating numbers from the uniform distribution, n = number of numbers to generat. 

runif(n=2,min=1,max=3)

colors[runif(1,1,3)]


colors[c(FALSE, FALSE, TRUE)]  # It shows only true values

greens <- colors == "green"  #returns which are true 

colors[greens]

```




```{r}
mytbl <- data.frame(colors, greens, yo=c("ye","YO", "Ya"), test = "test")
mytbl

#select rows
mytbl[1,]  #first row all columns

mytbl[-2:-3,]  # do not select 2nd and 3rd rows


mytbl[,-2:-3]  # do not select 2nd and 3rd colmns

mytbl[1:2,1:2]  #


mytbl[,"greens"]

mytbl[,c("colors","greens")]

mytbl$colors #  shorter way of getting the output is a vecktor

data.frame(mytbl$colors, mytbl$greens)


mytbl[mytbl$greens,]

mytbl[mytbl$yo=="YO", ] #select rows with YO

colnames(mytbl) =="greens"  #give true or false of which column name are "green""

mytbl[,colnames(mytbl)=="colors"]  #" same as above"

mytbl[,colnames(mytbl) > "colors"]   # show columns after colors, when sorted alphabetic order 




```






```{r}
#datapiplines
iris  %>%   # dataset, ctr+ shift+ m
  colnames() %>% 
  str_to_lower() ->
  newnames   #insert to newnames


```



```{r}
iris %>% # %>% takes previous insert into next 
  filter(Species=="setosa" , Sepal.Length + Sepal.Width < 10) %>%   #filter on conditions, , = and
  filter(Species=="setosa" | Sepal.Length + Sepal.Width < 10)  #filter on conditions, | = OR

```


```{r}
iris %>% # %>% takes previous insert into next 
  filter(Species=="setosa" , Sepal.Length + Sepal.Width < 10) %>%   #filter rows on conditions, , = and
  select(Sepal.Length,Sepal.Width, Species)  # selects columns
  

```


```{r}
iris %>% 
  select(Sepal.Length:Petal.Length, -Sepal.Length)  #select columns and unselect with -, last instruction tells it what to do finally
  
  
```


```{r}
iris %>% 
  select(starts_with('s'))  
```


```{r}
iris %>% 
  mutate(AjPetalLength = Petal.Length-0.1)  #adds a column , if written Petal.Length=Petal.Length-0.1 it would write it over
```

```{r}
iris %>% 
  mutate(Sepal.Area = Sepal.Width * Sepal.Length, Avg.Sepal.Area = mean(Sepal.Area)) # order sensitive
```



```{r}
iris %>% 
  group_by(Species)  %>%   #returns one value or row
  mutate(Sepal.Area = Sepal.Width * Sepal.Length, Avg.Sepal.Area = mean(Sepal.Area))   %>% 
  summarise(AvgSepal.Area = mean(Sepal.Length*Sepal.Width),
            unique(Avg.Sepal.Area)) #returns one value or row
```

```{r}
iris %>% 
  select(contains("Length"))

iris %>% 
  select(Petal.Width:Species, -contains("Length"))


```


```{r}
iris %>% 
  select_if(is.numeric)


iris %>% 
  select_if(~is.numeric(.) & n_distinct(.) > 30)  
# apply to each column two checks, se if numeric, dot (insert value here), if the also the value is unique values and larger than 30 select
```
[locke d mod](https://lockedata.co.uk/firebrandmodule2/dbplyr)



```{r, eval=FALSE} 
install.packages("datasauRus")
```

```{r}
library(ggplot2)
library(datasauRus)

ggplot(datasaurus_dozen, aes(x=x, y=y)) +  
  geom_point()

```



```{r}
ggplot(data = mtcars, aes(x = rownames(mtcars), y = mpg)) + geom_col() + theme(axis.text.x= element_text(angle=45, hjust=1))



```




```{r}


ggplot(data = mtcars, aes(x = rownames(mtcars), y = mpg, colour = as.factor(am))) + 
    geom_col(width = 0.5)


```



```{r}


ggplot(data = mtcars, aes(x = rownames(mtcars), y = mpg)) + 
    geom_col(width = 1)

```


```{r}

ggplot(data = iris, aes(x = Sepal.Width , y = Sepal.Length, color=Species)) + 
    geom_jitter(alpha=0.9)


```
#Barchart
```{r}
library(tidyverse)
datasaurus_dozen %>% 
  group_by(dataset) %>% 
  summarise(y=max(y)) %>% 
  ggplot(aes(x=dataset,y=y)) + geom_col()

```

#subplots with same size
```{r}
library(tidyverse)
ggplot(data = simpsons_paradox, aes(x = x, y = y)) + geom_point() + facet_wrap(~dataset, 
    nrow = 3)
```


# get rid of grey background
```{r}
library(tidyverse)
ggplot(simpsons_paradox, aes(x=x,y=y)) + 
  geom_point() + 
  facet_wrap(~dataset) + 
  theme_classic()
 # ggthemes :: theme_excel()
```

# GOODIE TOOLS!!
```{r, eval=FALSE} 
install.packages("devtools")
install.packages(c('xray', 'skimr', 'visdat', 'GGally'))
devtools::install_github('krupanss/IEDA')

```


# Gui or application for looking and working with the data before doing hard work
```{r}
library(tidyverse)

#IEDA::runIEDA()

#HRdata <-  read_csv('WA_Fn_UseC_HR_Employee_Attrition.csv')
#HRdata <-  WA_Fn_UseC_HR_Employee_Attrition

HRdata <- read_csv("~/ibm-hr-analytics-employee-attrition-performance/WA_Fn-UseC_-HR-Employee-Attrition.csv")


#skimr::skim(HRdata)

```

# skimr, 
## shows statistics, split out info, 
### Dataset1
```{r}
#library(tidyverse)

#HRdata <-  read_csv('WA_Fn_UseC_HR_Employee_Attrition.csv')
#HRdata <-  WA_Fn_UseC_HR_Employee_Attrition

skimr::skim(HRdata)

```



### Dataset2

```{r}
library(readr)
ecommerce_csv <- read_csv("ecommerce.csv")

"ecommerce.csv" %>% 
  read_csv(progress = FALSE) %>% 
  mutate(Description=str_remove_all(Description,"[^[:alnum:]]")) -> ecommerce

ecommerce %>% 
  skimr::skim()

#HRdata <-  read_csv('WA_Fn_UseC_HR_Employee_Attrition.csv')
#HRdata <-  WA_Fn_UseC_HR_Employee_Attrition



```


# Visdat
## shows the type of data  

```{r}

visdat::vis_dat(HRdata)

```

## visualize how much NaNs or missing data exists

```{r}

visdat::vis_miss(HRdata)
```


# Magic
```{r}

xray::anomalies(HRdata)


```


# Magic
```{r}

GGally::ggpairs(HRdata)
## too many plots for this example.

## reduce the columns being displayed
## these two lines of code produce the same plot matrix
pm <- GGally::ggpairs(HRdata, columns = c(1, 6, 2))
pm <- GGally::ggpairs(HRdata, columns = c("Age", "MonthlyIncome", "WorkLifeBalance"), columnLabels = c("Age", "MonthlyIncome", "WorkLifeBalance"))
pm

```




```{r, eval=FALSE}
install.packages("DataExplorer")

```

```{r}
DataExplorer::plot_bar(HRdata)   

```